A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-03-06 paper/paper1/ms.tex
text of constraining the mass loss rate

\section{1D Parker wind models and mass loss rates}
The dramatic leading tail of atmospheric escape suggests that more complicated 3D hydrodymaic simulations coupled with stellar wind \citep{2022ApJ...926..226M} may be needed to depict the geometry faithfully.  Here we explore one-dimensional (1D) models, which offer ease of interpretation and rapid calculation.  We employ the open-source \texttt{p-winds} code \citep{2022A&A...659A..62D}, which implements a transonic 1D Parker wind model with radiative transfer of the Helium 10830 triplet \citep{2018ApJ...855L..11O,2020A&A...636A..13L}.

@@

Equipped with these three SEDs of differing radiation hardness, we explored different mass loss rates and exosphere temperatures.  Figure \ref{fig:pwinds} shows one example model spectrum with $\dot{M} = 7\times10^{13}$ g/s, and $T_0=9100$~K.  The model spectrum exhibits approximately the same line width and depth as our HPF observations, making it a plausible solution.  The model-dependent line profile (\emph{i.e. width and depth}) is degenerate with XUV flux, $\dot{M}$, and $T_0$.  We were able to replicate the line profile with mass loss rates as high as \censorbox{X}, assuming \censorbox{Y}, or as low as \censorbox{Z} assuming \censorbox{W}.

@@

Figure XX shows the 1D wind velocity and density computed with \texttt{p-winds}, assuming the baseline planet properties, a characteristic exosphere temperature of $T\sim10^4$ K, and including tidal gravity.

@@

\subsection{Physical mechanisms for atmospheric escape}
\object{HAT-P-67 b} exhibits both dramatic atmospheric escape and significant radius inflation.  Ohmic dissipation stands out as a physical mechanism that could be responsible for both of these attributes.  \citet{2011ApJ...738....1B} showed that ohmic heating acts to inflate hot Jupiters, with low mass hot Jupiters overflowing their Roche lobes and leading to evaporation on Gyr timescales.  The Ohmic heating scenario requires only on a modest planetary magnetic field ($\sim$1 G) and enough photoionization to ionize some modest fraction of neutral metals.  The ``sweet spot'' for this phenomenon appears to prefer equilibrium temperatures in the range of $1500<T_\mathrm{eq}<2000$ K \citep{2011ApJ...738....1B}, well matched to HAT-P-67 b's $\sim$1900 K.  A plausible scaling law relationship approximate to an order-of-magnitude predicts an inflation timescales (Eq. 20 in \citet{2011ApJ...738....1B}):
\begin{equation}
    \tau_{\textit{infl}}\sim \left(\frac{0.01}{\epsilon} \right) \left(\frac{M}{M_{J}}\right)^2 \left(\frac{R_{J}}{R}\right)^3 \left(\frac{1500 \textrm{K}}{T_\mathrm{eff}}\right)^4 \textrm{Gyr}.
\end{equation}

@@

>>>>

\subsection{1D Parker wind models and mass loss rates}\label{pwinds}
The dramatic leading tail of atmospheric escape suggests that more complicated 3D hydrodymaic simulations coupled with stellar wind \citep{2022ApJ...926..226M} may be needed to depict the geometry faithfully.  Here we explore one-dimensional (1D) models, which offer ease of interpretation and rapid calculation.  We employ the open-source \texttt{p-winds} code \citep{2022A&A...659A..62D}, which implements a transonic 1D Parker wind model with radiative transfer of the Helium 10830 triplet \citep{2018ApJ...855L..11O,2020A&A...636A..13L}.

@@

Equipped with these three SEDs of differing radiation hardness, we explored different mass loss rates and exosphere temperatures.  Figure \ref{fig:pwinds} shows one example model spectrum with $\dot{M} = 7\times10^{13}$ g/s, and $T_0=9100$~K, with the high XUV spectrum ($L_x/L_\mathrm{bol}=10^{-4}$).  The model spectrum exhibits approximately the same line width and depth as our HPF observations, making it a conceivable solution.

At least a few shortcomings limit the applicability of this 1D Parker wind model.  First, the model-dependent line profile (\emph{i.e. width and depth}) is degenerate with XUV flux, $\dot{M}$, and $T_0$, and so a range of these parameters can be fine-tuned to obtain a large range of mass loss rates consistent with the data and our limited understanding of the XUV flux.  Second, the 1D model breaks down when attempting to explain the inherently 3D leading tail geometry.  We discuss these two and other model-dependent interpretations in Section \ref{secDiscuss}.

@@

@@

\subsection{Ohmic Dissipation and runaway escape}
\object{HAT-P-67 b} exhibits both dramatic atmospheric escape and significant radius inflation.  Ohmic dissipation stands out as a physical mechanism that could be responsible for both of these attributes.  \citet{2011ApJ...738....1B} showed that ohmic heating acts to inflate hot Jupiters, with low mass hot Jupiters overflowing their Roche lobes and leading to evaporation on Gyr timescales.  The Ohmic heating scenario requires only on a modest planetary magnetic field ($\sim$1 G) and enough photoionization to ionize some modest fraction of neutral metals.  The ``sweet spot'' for this phenomenon appears to prefer equilibrium temperatures in the range of $1500<T_\mathrm{eq}<2000$ K \citep{2011ApJ...738....1B}, well matched to HAT-P-67 b's $\sim$1900 K.  A plausible scaling law relationship approximate to an order-of-magnitude predicts an inflation timescales (Eq. 20 in \citet{2011ApJ...738....1B}):
\begin{equation}
    \tau_{\textit{infl}}\sim \left(\frac{0.01}{\epsilon} \right) \left(\frac{M}{M_{J}}\right)^2 \left(\frac{R_{J}}{R}\right)^3 \left(\frac{1500 \textrm{K}}{T_\mathrm{eff}}\right)^4 \textrm{Gyr}. \label{eqInflate}
\end{equation}

@@

\subsection{Constraining mass loss rate}
A 5 Myr inflation timescale would imply an instantaneous $\dot{M}_\mathrm{infl}\sim10^{15}$ g/s, about 50 times larger than the 1D Parker wind model mass loss rate of $\dot{M}_\mathrm{1D}\sim2\times10^{13}$. Here we enumerate all of uncertainties that go into assessing mass loss rates to show that some in-between mass loss rate may be plausible.

First, the scaling law arguments that produced Equation \ref{eqInflate} (Eq. 20 of \citet{2011ApJ...738....1B}) were only proposed as order of magnitude estimates not to be taken too faithfully.  So a $10\times$ higher inflation timescale of 50 Myr would yield $\dot{M}_\mathrm{infl}\sim10^{14}$ g/s, still a very large mass loss rate, but only a factor of 5 away from the baseline 1D model.  Order unity uncertainties in the Ohmic dissipation efficiency $\epsilon$ may also contribute.

As mentioned in Section \ref{pwinds}, inferring the mass loss rate from observations is strongly model dependent, as seen in \censorbox{cite e.g. S.V. papers}.  In particular, the uncertain XUV-to-NUV flux ratio introduces a large degeneracy in

the observations indicate a 3D geometry of the outflow-- a preference for escaping material to overtake the planet along the orbital path.  Once the The 1D Parker wind models for high mass loss rates
During the moment of the planet transit midcenter ($\phi=0.0$), the subsolar point on the planet dayside will have material be preferrentially removed from the line-of-sight
